--- 
knit: "bookdown::render_book"
title: "Making your codes re-usable using R packages "
subtitle: "GC Lunch and Learn: R packages 101"
author: ["Dmitry Gorodnichy and Joseph Stinziano"]
gitlab: https://gccode.ssc-spc.gc.ca/r4gc/gc-packages/packages101
date: "March -June 2021"
description: "The book is designed ...."
url: 'https://gccode.ssc-spc.gc.ca/r4gc/gc-packages/packages101'
github-repo: hadley/adv-r
site: bookdown::bookdown_site
documentclass: book
---

# Welcome {-}


This is the website for work-in-progress 2nd edition of __"[Advanced R](http://amzn.com/1466586966?tag=devtools-20)"__, a book in Chapman & Hall's R Series. The book is designed primarily for R users who want to improve their programming skills and understanding of the language.  It should also be useful for programmers coming to R from other languages, as it explains some of R's quirks and shows how some parts that seem horrible do have a positive side.


## Other books {-}

You may also be interested in:

* __"[R for Data Science](http://r4ds.had.co.nz/)"__ which introduces you to R 
  as a tool for doing data science, focussing on a consistent
  set of packages known as the tidyverse.

* __"[R Packages](http://r-pkgs.had.co.nz/)"__ which teaches you 
  how to make the most of R's fantastic package system.

<!--chapter:end:index-bookdown.Rmd-->


# Resources

## GC groups and repos

- Discussions: 
  - https://gccollab.ca/groups/profile/7391537/enuse-rfruse-r
  - https://gccollab.ca/groups/profile/7855030/enfriday-lunch-and-learn-r-meet-upsfr
- Codes:
  - Packages101 repo: https://gccode.ssc-spc.gc.ca/r4gc/gc-packages/packages101
  - Related codes and resources: https://gccode.ssc-spc.gc.ca/r4gc

## Related tutorials

- https://r-pkgs.org/
  - https://r-pkgs.org/vignettes.html
  
- https://fanwangecon.github.io/R4Econ/support/development/fs_packaging.pdf
- https://rstudio4edu.github.io/rstudio4edu-book/data-pkg.html - https://github.com/rstudio4edu/rstudio4edu-book 
- http://www.danieldsjoberg.com/writing-R-packages - github.com/ddsjoberg/writing-R-packages




# How to contribute

1. Anyone: Fork  Packages101 repo - modify - commit changes - push - submit request to merge

2. Members of r4gc group: 

A token is generated to allow you to push/pull Packages101 repo:  y5sVMrZPvskuFbeurZgH    
Use this line to push/pull from Packages101 repo:

    git clone --progress https://oauth2:tjxrg3GyUQJJDMaA6LfHA@gccode.ssc-spc.gc.ca/r4gc/gc-packages/packages101 r4gc_packages101


=======


# * R script to start:

```{r eval=FALSE, include=T}

library(devtools)

devtools::create("rCanada") # if you are  in /my_package
devtools::create("../../rCanada") # will it erase it,if I lready have it?
> #or usethis::create_package("../r4gc/packages/rCanada2")

library(roxygen2)
# library(testthis).<- we dont use this!Dont confuse it with library(testthat) ??

library(testthat) # do we need it?

library(usethis)
# https://github.com/r-lib/usethis

usethis::use_testthat()
# use_testthat()

use_news_md()

use_test("iviDT")

x <- 1
y <- 2
use_data(x, y)


use_vignette(name="my_vignettes") #
use_vignette(name="data_linking")


use_package("data.table")
# > use_package("data.table")
# √ Adding 'data.table' to Imports field in DESCRIPTION
# * Refer to functions with `data.table::fun()`


  
  

use_package("magrittr")
use_package("data.table")
use_package("lubridate", type="Imports")  # how about order ?! lubridate must be after data.table !
use_package("stringr")

use_package("IVIM")

# To update .Rd files in ./man, run:
devtools::document()
# Warning: The existing 'NAMESPACE' file was not generated by roxygen2, and will not be overwritten.
# So delete it, and then it will be created



```

## To be discussed

Q: 
When I Clean and Rebuid:
????! 
** testing if installed package can be loaded from final location

??????  C:/Users/gxd006/DOWNLO~1/R-40~1.2 /etc/Renviron.site ?? May 12 19:25:09 
Warning: replacing previous import 'data.table::month' by 'lubridate::month' when loading 'IVIM'



# * Setup

*This is what you already have: *  
- "Original"" folder, with no specific structure, 
with R and Rmd files that contain:
  a) functions that that you want to be re-used by others (and yourself many months later!).  
    - They are tested and test codes are included in `if(F)` or in separate .Rmd, or, even better, in interactive Shiny App (eg https://rCanada.shinyapp.io/covid)
    - Good idea to be have them in form that can be sourced: `source("caCovid.R"); source("iviBase.R")`
    
  b) functions and other codes that are not (yet) ready for re-use.

https://gccode.ssc-spc.gc.ca/r4gc/codes/tracking-covid-data
  - caCovid.R
  - iviBase.R
  -  ... common.R, plot.R, etc
  



*This is what you want to get:*
- Package folder (or several folders). 
  - GCCode/r4gc/packages/rCanada
  - GCCode/r4gc/packages/IVI 
https://gccode.ssc-spc.gc.ca/gorodnichy/rCanada

##  Ways to do it:

You can create it in two ways:

###  From New Project

Either way RStudio will initialize and launch your new project.
In the second way, it also puts  .gitignore

1.  In RStudio -> New Project -> R package (IVIM)
  
NB: there are many templates. Choose basic and create in folder /r4gc/packages

** Best way**
NB: If you have .R codes that are source-able, attached them from the menu!

It will generate less details (e.g in DESCRIPTION) than second method below.


###  Other methods and advice from Hadley:


https://r-pkgs.org/workflows101.html
  
Call usethis::create_package("path/to/package/pkgname").

In RStudio, do File > New Project > New Directory > R Package. This ultimately calls usethis::create_package(), so really there’s just one way.

Don’t use package.skeleton() to create a package. Because this function comes with R, you might be tempted to use it, but it creates a package that immediately throws errors with R CMD build. 


  
  
###   From script

2. Run from any R, Rmd window or R console the command below


NB: output from 

```
> devtools::create("../r4gc/packages/rCanada")
> #or usethis::create_package("../r4gc/packages/rCanada2")

New project 'rCanada' is nested inside an existing project '../r4gc/packages/', which is rarely a good idea.
If this is unexpected, the here package has a function, `here::dr_here()` that reveals why '../r4gc/packages/' is regarded as a project.
Do you want to create anyway?

1: Yes
2: Absolutely not
3: Nope

Selection: 1
√ Creating '../r4gc/packages/rCanada/'
√ Setting active project to 'C:/Users/gxd006/Downloads/_CODES/GCCode/r4gc/packages/rCanada'
√ Creating 'R/'
√ Writing 'DESCRIPTION'
Package: rCanada
Title: What the Package Does (One Line, Title Case)
Version: 0.0.0.9000
Authors@R (parsed):
    * First Last <first.last@example.com> [aut, cre] (YOUR-ORCID-ID)
Description: What the package does (one paragraph).
License: `use_mit_license()`, `use_gpl3_license()` or friends to
    pick a license
Encoding: UTF-8
LazyData: true
Roxygen: list(markdown = TRUE)
RoxygenNote: 7.1.1
√ Writing 'NAMESPACE'
√ Writing 'rCanada.Rproj'
√ Adding '^rCanada\\.Rproj$' to '.Rbuildignore'
√ Adding '.Rproj.user' to '.gitignore'
√ Adding '^\\.Rproj\\.user$' to '.Rbuildignore'
√ Opening 'C:/Users/gxd006/Downloads/_CODES/GCCode/r4gc/packages/rCanada/' in new RStudio session
√ Setting active project to '<no active project>'
```



# * Overall Workflow



## .. Copy needed codes from MY_CODES to R directory

- Copy a function
```
dt.replaceAwithB <- function(dt, col, a, b) {
  dt[get(col)==a, (col):=b]; 
}
```
- Insert roxigen skeleton (from magic wand menu) and add description
NB: you may need to manually insert @ import and @ export


###

no need to Load All?
Always start from "Clean and rebuild"
Then Check 



# * .Rbuildignore
^.*\.Rproj$
^\.Rproj\.user$
MY_CODES\ incorrect
^MY_CODES$ correct
^MY_DATASETS$
^LICENSE\.md$



# * License: `use_mit_license()`
License: MIT + file LICENSE




# * DESCRIPTION

# * NAMESPACE

https://r-pkgs.org/namespace.html

```
use_package("data.table")
# > use_package("data.table")
# √ Adding 'data.table' to Imports field in DESCRIPTION
# * Refer to functions with `data.table::fun()`
```

However this does notchange NAMESPACE ! Who changes is it?
is that devtools::document() ???

Generated by roxygen2: do not edit by hand
```
export()
export(addDerivatives)
export(extractMostInfectedToday)
export(readCovidUofT.csv)
importFrom(lubridate,dmy)
importFrom(stringr,str_replace)
```
I deleted NAMESPACE so that roxigen can generate it!
then added manually there . Not sure that's theway to do it !
```
import(data.table)
import(ggplot2)
import(lubridate)
import(magrittr)
import(IVIM)
```

It Worked !

  If you are using just a few functions from another package, the recommended option is to note the package name in the Imports: field of the DESCRIPTION file and call the function(s) explicitly using ::, e.g., pkg::fun(). Alternatively, though no longer recommended due to its poorer readability, use @importFrom, e.g., @importFrom pgk fun, and call the function(s) without ::.



# * Examples and tests

## .. In main file  in /R folder

```
#' @examples
```

Some could be such :
```
#' \dontrun{}
```

```{r create-documentation, eval=FALSE, include=FALSE}
devtools::document()

```

## .. In /tests folder


```
> use_testthat()

√ Setting active project to 'C:/Users/gxd006/Downloads/_CODES/GCCode/r4gc/packages/IVIM'
√ Adding 'testthat' to Suggests field in DESCRIPTION
√ Setting Config/testthat/edition field in DESCRIPTION to '3'
√ Creating 'tests/testthat/'
√ Writing 'tests/testthat.R'
* Call `use_test()` to initialize a basic test file and open it for editing.
Warning messages:
1: In readLines(f, n) :
  incomplete final line found on 'C:/Users/gxd006/Downloads/_CODES/GCCode/r4gc/packages/IVIM/DESCRIPTION'
...
  incomplete final line found on 'C:/Users/gxd006/Downloads/_CODES/GCCode/r4gc/packages/IVIM/DESCRIPTION'
> use_vignette()
Error in check_vignette_name(name) : 
  argument "name" is missing, with no default
```

# .. in MY_CODES

Provide as .R, Rmd or shiny


# * testthat


https://testthat.r-lib.org/




# * Documentation


devtools::document()

https://kbroman.org/pkg_primer/pages/docs.html

```
#' For more details see the help vignette:
#' \code{vignette("help", package = "mypkg")}

or 

\href{../doc/help.html}{\code{vignette("help", package = "mypkg")}}

```


# * Vignettes


```  
> use_vignette(name="my_vignettes")

√ Adding 'knitr' to Suggests field in DESCRIPTION
√ Setting VignetteBuilder field in DESCRIPTION to 'knitr'
√ Adding 'inst/doc' to '.gitignore'
√ Creating 'vignettes/'
√ Adding '*.html', '*.R' to 'vignettes/.gitignore'
√ Adding 'rmarkdown' to Suggests field in DESCRIPTION
√ Writing 'vignettes/my_vignettes.Rmd'
* Modify 'vignettes/my_vignettes.Rmd'
```

I later renamed my first vignette to intro.Rmd - manually



https://bookdown.org/yihui/rmarkdown-cookbook/package-vignette.html




# Packaging and publishing w. `pkgdown`

https://fanwangecon.github.io/R4Econ/support/development/fs_packaging.pdf


# Appendix: Various possible Errors and problems



### Errors in using usethis::use_testthat() 


Problem:   
```

> usethis::use_vignette(name="howtoCovid")
Error in read.dcf(con) : 
  Line starting 'analyze and visualiz ...' is malformed!
> 
```
Solution:    
https://stackoverflow.com/questions/59303030/devtools-error-in-read-dcfpath-desc-line-starting-this-corresponds-to

Corrections from https://gccode.ssc-spc.gc.ca/r4gc/gc-packages/IVIM/-/tree/useR-dev/



# (optional) Rtools

NBL This is not required

https://cran.r-project.org/bin/windows/Rtools/

http://web.mit.edu/insong/www/pdf/rpackage_instructions.pdf

<!--chapter:end:../packages101.Rmd-->

